### 2013-05-22 paper.tex
New paper

\documentclass[10pt,conference,a4paper]{IEEEtran}

%import packages here
\usepackage{amsmath}
\everymath{\displaystyle}
\usepackage{graphicx}
\usepackage{caption}
\graphicspath{{Images/}}

\title{Internet Attribute Certificate Profile for Authorization - An Analyses}
\author{Ruud Verbij}
%begin document
\begin{document}
\maketitle

\begin{abstract}
ABSTRACT
\end{abstract}

\begin{IEEEkeywords}
Keywords
\end{IEEEkeywords}

\label{Introduction}
\section{Introduction}
Below is a list of 8 RFC assignments. Read and study the RFC (skip the very technical stuff such as ASN.1, focus on cryptographic and PKI matters). Then give a summary of the RFC on a high level, addressing the following issues:
\begin{enumerate}
	\item How does this RFC fit in the entire PKI picture? Who is expected to use this RFC,what for, and how?
	\item What are the security requirements outlined in this RFC? What security problems does it attempt to solve? What (cryptographic) attacks does it defend against?
	\item Give an overview of the solution that the RFC proposes. How is cryptography used in meeting the security requirements, solving the security problems, defending against the attacks?
\end{enumerate}
\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv,literature}

\end{document}